Area maximisation in continuous time

### 2013-11-08 Notes/ContTime.tex
laplacian envelope written up

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{mydefs.tex} with its contents
%       or include mydefs.tex in zip/tar file
%    Replace \input{newsymb.tex} with its contents
%       or include newsymb.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
%
% >> Only for drafts! <<
\usepackage[notref,notcite]{showkeys}
% ----------------------------------------------------------------
%\numberwithin{equation}{section}
%\renewcommand{\baselinestretch}{1.5}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\inv}{^{-1}}
\newcommand{\dg}{^\mathrm{dg}}
\newcommand{\trans}{^\mathrm{T}}
\newcommand{\I}{\mathbf{I}}
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MM}{\mathbf{Q}}
%equilibrium distribution
\newcommand{\eq}{\mathbf{p}^\infty}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fund}{\mathbf{Z}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\Lb}{\boldsymbol{\Lambda}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\uv}{\mathbf{u}}
\newcommand{\vv}{\mathbf{v}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\enc}{\mathbf{q}}
\newcommand{\frg}{\W^{\mathrm{F}}}
\newcommand{\F}{\boldsymbol{\Phi}}
\DeclareMathOperator{\SNR}{SNR}
\newcommand{\CS}{\mathcal{S}}
\newcommand{\CA}{\mathcal{A}}
\newcommand{\CB}{\mathcal{B}}
\newcommand{\comp}{^\mathrm{c}}
% ----------------------------------------------------------------
\input{sl_theorems_preamble.tex}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{Area maximisation in continuous time}
%
% Author List:
%
\author{Subhaneil Lahiri
%
}

@@

Now consider the Laplace transform of the SNR curve \eqref{eq:SNRcurve}
%
\begin{equation}\label{eq:lsnr}
\begin{aligned}
  A(s) &= \int_0^{\infty}\dr t \e^{-st}\SNR(t) \\
   &= \int_0^{\infty}\dr t\, \sqrt{N}(2f^+f^-) \eq \enc \e^{(r\frg-s)t} \w \\
   &= \frac{\sqrt{N}(2f^+f^-)}{r} \eq \enc \prn{s - r\frg}\inv \w.
\end{aligned}
\end{equation}
%
Note that $A(0)=A$, the area, and $\lim_{s\to\infty} \brc{sA(s)} = \SNR(0)$, the initial SNR.

Thanks to \eqref{eq:alfund}, we can replace $\mathbf{G}(s) \to \fund(s)$, as $\enc\onev=0$.
%
\begin{equation}\label{eq:larea}
\begin{aligned}
  \hat{A}(s) = \frac{rA(s)}{\sqrt{N}(2f^+f^-)} &=  \eq \enc \fund(s) \w \\
    &= \sum_{ijk} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \eq_k \w_k \\
    &= 2\sum_{ij}  \eq_i \enc_{ij} (\eta^+_{i}(s) - \eta^+_{j}(s)).
\end{aligned}
\end{equation}
%
In analogy with \eqref{eq:areacoeffs}, we define
%
\begin{equation}\label{eq:lareacoeffs}
  \begin{aligned}
    c_k(s) &= \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \\
    a_i(s) &= \sum_j \eq_i \enc_{ij} (\eta^+_{i}(s) - \eta^+_{j}(s)),\\
    \implies
    \hat{A}(s) &= \sum_k c_k(s) \eq_k \w_k
      = 2\sum_i a_i(s).
  \end{aligned}
\end{equation}
%

@@

Using the same approach as the area bound, we find that the derivatives are
%
\begin{equation}\label{eq:lAdiff}
  \pdiff{\hat{A}(s)}{\W^\pm_{gh}} = 2rf^\pm \eq_g \brk{ \sum_i a_i(s) (\fptb_{gi}(0)-\fptb_{hi}(0))
     +c_g(s) (\eta^+_g(s) - \eta^+_h(s)) } \pm 2\eq_g (\eta^+_g(s) - \eta^+_h(s)).
\end{equation}
%
This is exactly the same as \eqref{eq:derivA}.
It follows that, using exactly the same perturbations, that the Laplace transform of the SNR curve for \emph{any} model is lower than that of some model with the serial topology having the same equilibrium distribution.

>>>>

% -*- TeX -*- -*- UK -*-
% ----------------------------------------------------------------
% arXiv Paper ************************************************
%
% Subhaneil Lahiri's template
%
% Before submitting:
%    Comment out hyperref
%    Comment out showkeys
%    Replace \input{mydefs.tex} with its contents
%       or include mydefs.tex in zip/tar file
%    Replace \input{newsymb.tex} with its contents
%       or include newsymb.tex in zip/tar file
%    Put this file, the .bbl file, any picture or
%       other additional files and natbib.sty
%       file in a zip/tar file
%
% **** -----------------------------------------------------------
\documentclass[12pt]{article}
%Preamble:
\usepackage{a4wide}
\input{sl_preamble.tex}
%
% >> Only for drafts! <<
\usepackage[notref,notcite]{showkeys}
% ----------------------------------------------------------------
%\numberwithin{equation}{section}
%\renewcommand{\baselinestretch}{1.5}
% ----------------------------------------------------------------
% New commands etc.
\input{sl_definitions.tex}
\input{sl_symbols.tex}
%matrices
\newcommand{\inv}{^{-1}}
\newcommand{\dg}{^\mathrm{dg}}
\newcommand{\trans}{^\mathrm{T}}
\newcommand{\I}{\mathbf{I}}
%vec of ones
\newcommand{\onev}{\mathbf{e}}
%mat of ones
\newcommand{\onem}{\mathbf{E}}
%Markov matrix
\newcommand{\MM}{\mathbf{Q}}
%equilibrium distribution
\newcommand{\eq}{\mathbf{p}^\infty}
%first passage times
\newcommand{\fpt}{\mathbf{T}}
%off-diag first passage times
\newcommand{\fptb}{\overline{\fpt}}
%fundamental matrix
\newcommand{\fund}{\mathbf{Z}}
%other symbols
\newcommand{\Pb}{\mathbf{P}}
\newcommand{\D}{\mathbf{D}}
\newcommand{\pib}{\boldsymbol{\pi}}
\newcommand{\Lb}{\boldsymbol{\Lambda}}
\newcommand{\w}{\mathbf{w}}
\newcommand{\uv}{\mathbf{u}}
\newcommand{\vv}{\mathbf{v}}
\newcommand{\W}{\mathbf{W}}
\newcommand{\M}{\mathbf{M}}
\newcommand{\enc}{\mathbf{q}}
\newcommand{\frg}{\W^{\mathrm{F}}}
\newcommand{\F}{\boldsymbol{\Phi}}
\DeclareMathOperator{\SNR}{SNR}
\DeclareMathOperator{\snr}{SNR}
\newcommand{\CI}{\mathcal{I}}
\newcommand{\CS}{\mathcal{S}}
\newcommand{\CA}{\mathcal{A}}
\newcommand{\CB}{\mathcal{B}}
\newcommand{\comp}{^\mathrm{c}}
% ----------------------------------------------------------------
\input{sl_theorems_preamble.tex}
% ----------------------------------------------------------------
%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Title info:
\title{Area maximisation in continuous time}
%
% Author List:
%
\author{Subhaneil Lahiri
%
}

@@

Now consider the Laplace transform of the SNR curve \eqref{eq:SNRcurve}
%
\begin{equation}\label{eq:lsnr}
\begin{aligned}
  A(s) &= \int_0^{\infty}\dr t \e^{-st}\SNR(t) \\
   &= \int_0^{\infty}\dr t\, \sqrt{N}(2f^+f^-) \eq \enc \e^{(r\frg-s)t} \w \\
   &= \frac{\sqrt{N}(2f^+f^-)}{r} \eq \enc \prn{s - r\frg}\inv \w.
\end{aligned}
\end{equation}
%
This expression is ill-behaved at $s=0$.
Thanks to \eqref{eq:alfund}, we can solve this by the replacement $\mathbf{G}(s) \to \fund(s)$, as $\enc\onev=0$.
%
\begin{equation}\label{eq:larea}
\begin{aligned}
  \hat{A}(s) = \frac{rA(s)}{\sqrt{N}(2f^+f^-)} &=  \eq \enc \fund(s) \w \\
    &= \sum_{ijk} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \eq_k \w_k \\
    &= 2\sum_{ij}  \eq_i \enc_{ij} (\eta^+_{i}(s) - \eta^+_{j}(s)).
\end{aligned}
\end{equation}
%
Note that $A(0)=A$, the area, and $\lim_{s\to\infty} \brc{sA(s)} = \SNR(0)$, the initial SNR.

In analogy with \eqref{eq:areacoeffs}, we define
%
\begin{equation}\label{eq:lareacoeffs}
  \begin{aligned}
    c_k(s) &= \sum_{ij} \eq_i \enc_{ij} \prn{\fptb_{ik}(s)-\fptb_{jk}(s)} \\
    a_i(s) &= \sum_j \eq_i \enc_{ij} (\eta^+_{i}(s) - \eta^+_{j}(s)),\\
    \implies
    \hat{A}(s) &= \sum_k c_k(s) \eq_k \w_k
      = 2\sum_i a_i(s).
  \end{aligned}
\end{equation}
%

@@

Using the same approach as the area bound, we find that the derivatives are
%
\begin{equation}\label{eq:lAdiff}
\begin{aligned}
  \pdiff{\hat{A}(s)}{\W^\pm_{gh}} 
     =&\, 2rf^\pm \eq_g \brk{ \sum_i a_i(s) (\fptb_{gi}(0)-\fptb_{hi}(0))
     +c_g(s) (\eta^+_g(s) - \eta^+_h(s)) } \\
     &\pm 2\eq_g (\eta^+_g(s) - \eta^+_h(s)).
\end{aligned}
\end{equation}
%
This is exactly the same as \eqref{eq:derivA}.
It follows that, using exactly the same perturbations, that the Laplace transform of the SNR curve for \emph{any} model is lower than that of a certain model with the serial topology having the same equilibrium distribution.

\subsection{Envelope}\label{sec:env}

Hopefully we can prove an upper bound on the Laplace transform of the memory curve:
%
\begin{equation}\label{eq:lbound}
  A(s) \leq \sqrt{N}B(s) = \frac{\sqrt{N}}{C(s)}.
\end{equation}
%
We will assume that $C(s)$ is a concave function.
We already know that $C(0)=\frac{1}{M-1}$ and $C(s) \to s$ as $s\to\infty$.

In the eigenmode decomposition, this reads
%
\begin{equation}\label{eq:lboundeig}
  \sum_a \frac{\CI_a}{q_a+s} \leq B(s).
\end{equation}
%
Now let's maximise $\snr(t_0)$ with respect to $\brc{\CI_a,q_a}$.
Consider the Lagrangian
%
\begin{equation}\label{eq:envlagrange}
  \CL = \sum_a \CI_a \e^{-q_a t_0} +\intd[_0^\infty]{s} \mu(s)\brk{B(s)-\sum_a\frac{\CI_a}{q_a+s}}.
\end{equation}
%
Necessary conditions for a maximum are
%
\begin{alignat}{3}
  \pdiff{\CL}{\CI_a} &= \e^{-q_a t_0} - \intd{s} \frac{\mu(s)}{q_a+s} = 0, \label{eq:envI}\\
  \pdiff{\CL}{q_a} &= \CI_a\brk{-t_0\e^{-q_a t_0} + \intd{s} \frac{\mu(s)}{(q_a+s)^2}} = 0, \label{eq:envq}\\
  \mu(s)\fdf{\CL}{\mu(s)} &= \mu(s)\brk{B(s)-\sum_a\frac{\CI_a}{q_a+s}} = 0, \label{eq:envKT}\\
  \mu(s) &\geq 0, \label{eq:envKTm}\\
  \fdf{\CL}{\mu(s)} &= B(s)-\sum_a\frac{\CI_a}{q_a+s} \geq 0 \label{eq:envKTbnd}.
\end{alignat}
%
If we assume $\CI_a \neq0$ for some $a$, then \eqref{eq:envI} and \eqref{eq:envq} read
%
\begin{equation}\label{eq:envnumexp}
\begin{aligned}
  \e^{-q_a t_0} &= \intd{s} \frac{\mu(s)}{q_a+s}, \\
  t_0\e^{-q_a t_0} &= \intd{s} \frac{\mu(s)}{(q_a+s)^2}.
\end{aligned}
\end{equation}
%
The second of these is the derivative of the first with respect to $q$, therefore the two curves in the first must touch each other tangentially at $q_a$.
As both curves are decreasing convex functions, it seems unlikely that this could happen at multiple values of $q$.
We will assume that there is a single non-zero $\CI_a=\CI$, with the corresponding $q_a=q$, in the following.

For \eqref{eq:envI} to have a solution, $\mu(s)$ must be nonzero for at least one value of $s$.
Then, by \eqref{eq:envKT}, we must saturate \eqref{eq:envKTbnd} at those values of $s$.
If there were more than two such values, $\CI$ and $q$ would be overconstrained.
Assume there are two such values, $s=s_0$ and $s=s_1$.
%
\begin{equation}\label{eq:sattwice}
\begin{aligned}
  \CI &= B(s_0)(q+s_0) = B(s_1)(q+s_1) \qquad
  \implies q &= \frac{s_0 B(s_0) - s_1 B(s_1)}{B(s_1)-B(s_0)}.
\end{aligned}
\end{equation}
%
Now the bound \eqref{eq:envKTbnd} reads
%
\begin{equation}\label{eq:satwicebnd}
\begin{aligned}
  B(s) \geq \frac{B(s_0)(q+s_0)}{q+s} &= \frac{s_0-s_1}{(s-s_1)C(s_0)+(s_0-s)C(s_1)}, \\
  xC(s_0)+(1-x)C(s_1) &\geq C(xs_0+(1-x)s_1),
\end{aligned}
\end{equation}
%
where we substituted $s=xs_0+(1-x)s_1$.
As we assumed that $C(s)$ is concave, this is impossible.

Therefore the bound must be saturated exactly once and $\mu(s)=\mu_0\delta(s-s_0)$.
Now \eqref{eq:envI}, \eqref{eq:envq} and \eqref{eq:envKT} read
%
\begin{equation}\label{eq:satonce}
  \mu_0 = (q+s_0)\e^{-qt_0},
  \qquad
  (q+s_0)t_0 = 1,
  \qquad
  \CI = \frac{B(s_0)}{t_0}.
\end{equation}
%
This satisfies \eqref{eq:envKTm}.
Now \eqref{eq:envKTbnd} tells us
%
\begin{equation}\label{eq:satoncebnd}
\begin{aligned}
  B(s) &\geq \frac{B(s_0)}{(q+s)t_0}, 
  \qquad\text{or}\qquad
  1 + (s-s_0)t_0 &\geq \frac{C(s)}{C(s_0)}.
\end{aligned}
\end{equation}
%
As the two sides are equal at $s=s_0$, it is necessary that they touch tangentially for this bound to be true elsewhere.
The linearity of the LHS and concavity of the RHS ensure that this is sufficient for the bound to be satisfied for all $s$.
This condition is
%
\begin{equation}\label{eq:envs0t0}
  t_0 = \frac{C'(s_0)}{C(s_0)}.
\end{equation}
%

The resulting SNR curve and envelope are
%
\begin{equation}\label{eq:env}
  \snr(t) = \frac{\e^{(s_0-1/t_0)t}}{C'(s_0)},
  \qquad
  \snr(t_0) = \frac{\e^{s_0t_0-1}}{C'(s_0)},
\end{equation}
%
where $s_0$ is determined as a function of $t_0$ by \eqref{eq:envs0t0}.
